---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ABwaterwells

<!-- badges: start -->

[![R-CMD-check](https://github.com/stevenpawley/ABwaterwells/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/stevenpawley/ABwaterwells/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/stevenpawley/ABwaterwells/graph/badge.svg)](https://app.codecov.io/gh/stevenpawley/ABwaterwells)
<!-- badges: end -->

**ABwaterwells** provides an R interface to the [Alberta Water Well Information Database](https://www.alberta.ca/alberta-water-well-information-database) (AWWID) via its OData web service. Large tables are downloaded in parallel chunks and automatically converted from imperial to metric units.

## Installation

```r
# install.packages("devtools")
devtools::install_github("stevenpawley/ABwaterwells")
```

## Getting started

`awwid_connect()` fetches the list of available tables once and returns a
connection object with one named field per table.

```{r connect}
library(ABwaterwells)

con <- awwid_connect()
con
```

## Accessing tables by name

Pass any field from the connection object directly to `awwid_tbl()`. Most IDEs
will tab-complete the available table names, without the need to type or remember them
as characters.

```{r named-methods}
wells <- awwid_tbl(
  con$wells,
  select = c("gicwellid", "wellid", "longitude", "latitude"),
  filter = "gicwellid gt 40000 and gicwellid lt 41000"
)

reports <- awwid_tbl(
  con$wellreports,
  select = c("wellid", "wellreportid", "totaldepthdrilled"),
  filter = "wellreportid gt 40000 and wellreportid lt 41000"
)

lithologies <- awwid_tbl(
  con$lithologies,
  filter = "wellreportid gt 40000 and wellreportid lt 41000"
)

head(lithologies)
```

## Looping over multiple tables

For programmatic use such as iterating over a variable set of table names, pass a
character string to `awwid_tbl()` directly, or index into the connection object
with `[[`:

```{r loop-tables, eval = FALSE}
tables_to_download <- c("wells", "wellreports", "lithologies")

results <- lapply(tables_to_download, function(tbl) {
  awwid_tbl(con[[tbl]], top = 100)
})
names(results) <- tables_to_download
```

## Converting to metric units

Data in AWWID is stored in imperial units. Call `metricate()` to convert a
table automatically:

```{r metricate}
reports_df    <- metricate(reports)
lithologies_df <- metricate(lithologies)

head(lithologies_df)
```

## Predefined queries

Several multi-table queries cover common workflows. Each accepts already-
downloaded, metricated tables as arguments, so you control when downloads happen
and can reuse data across calls.

```{r queries}
lithologs <- query_lithologs(wells, reports_df, lithologies_df)
lithologs
```

```{r queries-screens, eval = FALSE}
screens      <- awwid_tbl(con$screens,
                          filter = "wellreportid gt 40000 and wellreportid lt 41000")
perforations <- awwid_tbl(con$perforations,
                          filter = "wellreportid gt 40000 and wellreportid lt 41000")

query_screens(
  wells        = wells,
  wells_reports = reports_df,
  screens      = metricate(screens),
  perforations = metricate(perforations),
  .aggregate   = TRUE,
  .assumed_top = 5
)
```
