---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ABwaterwells

<!-- badges: start -->

<!-- badges: end -->

**ABwaterwells** provides an R interface to the [Alberta Water Well Information Database](https://www.alberta.ca/alberta-water-well-information-database) (AWWID) via its OData web service. Large tables are downloaded in parallel chunks and automatically converted from imperial to metric units.

## Installation

```r
# install.packages("devtools")
devtools::install_github("stevenpawley/ABwaterwells")
```

## Getting started

Create a connection object. This fetches the list of available tables from the service.

```{r connect}
library(ABwaterwells)

con <- AwwidQuery$new()
```

## Accessing tables by name

Each AWWID table has a dedicated method on the connection object. Use your IDE's tab completion to discover available tables — no need to remember or type table names as strings.

```{r named-methods}
# named methods accept filter, select, and top arguments
wells <- con$wells(
  select = c("gicwellid", "wellid", "longitude", "latitude"),
  filter = "gicwellid in (40000, 40001, 40002)"
)

reports <- con$wellreports(
  select = c("wellid", "wellreportid", "totaldepthdrilled"),
  filter = "wellreportid in (40000, 40001, 40002)"
)

# downloaded data is stored in the $data field
head(reports$data)
```

Data in AWWID is stored in imperial units. Call `$metricate()` to convert a table to metric:

```{r metricate}
wells_df   <- wells$metricate()
reports_df <- reports$metricate()

head(reports_df)
```

## Looping over multiple tables

To loop over a set of table names programmatically, use `$request()` directly. The `$tables` field lists every table available on the service.

```{r list-tables}
con$tables
```

```{r loop-tables, eval = FALSE}
tables_to_download <- c("wells", "wellreports", "lithologies")

results <- lapply(tables_to_download, function(tbl) {
  con$request(tbl, top = 100)$metricate()
})
names(results) <- tables_to_download
```

## Caching

Repeated calls with the same arguments return the cached result immediately. Use `$clear_cache()` to force a fresh download.

```{r caching}
# second call returns instantly from cache
con$wells(
  select = c("gicwellid", "wellid", "longitude", "latitude"),
  filter = "gicwellid in (40000, 40001, 40002)"
)
```

## Predefined queries

Several multi-table queries cover common workflows. Caching means shared tables (e.g. `wells`) are only downloaded once across queries.

```{r queries, eval = FALSE}
lithologs  <- con$query_lithologs()
waterlevel <- con$query_water_level()
```
